Species motif participation provides unique information about species risk of extinction

### 2022-12-06 manuscript/reorder_main_text.tex
formatted abstract for JAE

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[numbers]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{xr}

@@

\author{Alyssa R. Cirtwill$^{1*\dagger}$, Anna \r{A}kesson$^{2*}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{2}$} 
\date{
\small$^1$Spatial Foodweb Ecology Group\\
Research Centre for Ecological Change\\
Organismal and Evolutionary Biology Research Programm\\
Faculty of Biological and Environmental Sciences\\
University of Helsinki, Helsinki, Finland\\
\medskip
\small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University, Link\"{o}ping, Sweden\\
\medskip
\small$^3$ BioFrontiers Institute\\
University of Colorado, Boulder, Boulder, U.S.A.\\
\medskip
* Joint first authors\\
\medskip
$^\dagger$ Corresponding author\\
}

\begin{document} 
\maketitle 
\linenumbers
\raggedright

@@

\section*{Abstract}

    Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.
    Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance.
    We are interested in whether a richer set of species roles --based on participation in three-species motifs-- is also related to persistence and how these motif-based roles relate to simpler measures.
    We show that a consumer species' participation in different motifs is related to its probability of persistence. 
    Similarly, the overall motif profile of a network is related to the average persistence of consumer species.
    Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance, while more frequent participation in the omnivory motif was associated with higher species-level probabilities of persistence.
    Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level.
    Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.

\begin{spacing}{2.0}

\clearpage
\section*{Introduction} %New

@@

\bibliographystyle{vancouver} 
\bibliography{anna_bib_new} % Abbreviate journal titles.

>>>>

\documentclass[12pt]{article} 
\usepackage{amsmath} 
\usepackage[dvips]{graphicx}
\usepackage{multirow} 
\usepackage{geometry} 
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage[labelfont=bf]{caption} 
\usepackage{setspace}
\usepackage[running]{lineno} 
% \usepackage[numbers,sort]{natbib}
\usepackage[sort]{natbib} 
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{xr}
\usepackage{indentfirst}

@@

\author{Alyssa R. Cirtwill$^{1*\dagger}$, Anna \r{A}kesson$^{2*}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{2}$} 
\date{
\small$^1$Spatial Foodweb Ecology Group\\
Research Centre for Ecological Change\\
Organismal and Evolutionary Biology Research Programm\\
Faculty of Biological and Environmental Sciences\\
University of Helsinki, Helsinki, Finland\\
\medskip
\small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University, Link\"{o}ping, Sweden\\
\medskip
\small$^3$ BioFrontiers Institute\\
University of Colorado, Boulder, Boulder, U.S.A.\\
\medskip
* Joint first authors\\
\medskip
$^\dagger$ Corresponding author email: alyssa.cirtwill@gmail.com\\
}

\begin{document} 
\maketitle 
\raggedright

@@

\section*{Running headline}
Motif participation describes extinction risk

\section*{Summary}

    \indent{1. Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
    2. Coarse measures of a species' structural role, like in-degree and trophic level, are known to affect a consumer's vulnerability to bottom-up disturbance. \\
    3. We are interested in whether a richer set of species roles --based on participation in three-species motifs-- is also related to persistence and how these motif-based roles relate to simpler measures. \\
    4. We show that a consumer species' participation in different motifs is related to its probability of persistence. \\ 
    5. Similarly, the overall motif profile of a network is related to the average persistence of consumer species. \\
    6. Notably, we found that networks containing a higher proportion of the `omnivory' motif tend to have lower mean persistence following disturbance, while more frequent participation in the omnivory motif was associated with higher species-level probabilities of persistence. \\
    7. Additionally, we find that relationships between motif participation and persistence can synthesize the information provided by in-degree and trophic level. \\
    8. Motif participation therefore captures important trends in species persistence and provides a rich description of species' structural roles in their communities.

\section*{Key-words}
    food web, degree, trophic level, omnivory, competition

\begin{spacing}{2.0}
\linenumbers

\section*{Introduction} %New

@@

\bibliographystyle{jae} 
\bibliography{anna_bib_new} % Abbreviate journal titles.
